---
title: "W47 Learning Journal"
author: "Mette Emily Kendon"
Date: 23/11/2020
---

#Thoughts/intentions:  
My thoughts on how it was going to be working with R and RStudio was very much based on the R-tutorial. I had read the R-tutorial before the hands-on-session and I therefore had a good indication of how to move around in R. That was also why I didn’t feel lost about this assignment. I knew that I was going to use R for my final project and therefor my intention with this assignment was to fully understand how to use R and RStudio, in order to learn the different possibilities when working with this kind of software. 

#Exercises:
* Use R to figure out how many elements in the vector below are greater than 2. (You need to filter out the NAs first). rooms <- c(1, 2, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, NA)

I started by opening RStudio and made sure that I was in the directory I had created at the hands-on-session by typing:

```{r}
getwd()
```

This command asks the computer "where am I now"? It was important to be in the right directory because I had saved my script from the hands-on-session, and I knew that this could give me some inspiration on how to solve the four exercises.

I opened a new script by pressing the + in the left-hand corner of the window and chose R Markdown. This is a good feature in RStudio, because it allows you to document all of your steps. In this window I typed:

```{r}
rooms <- c(1, 2, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, NA)
```

This command assigned the vector () to the object named rooms. 
Then I typed:

```{r}
rooms_clean <- na.omit(rooms)
```

This command removed the NA’s and created a new vector containing only the number of rooms in each household. This new vector was assigned to the object named rooms_clean. 
Then I typed:

```{r}
rooms_above_2 <- rooms_clean[rooms_clean > 2]
```

This command created a new vector containing all of the rooms of the households from the object rooms_clean that had more than two rooms and placed them in a new object called rooms_above_2. Then I used one last command to ask RStudio how many of the households had more than 2 rooms by typing:

```{r}
length(rooms_above_2)
```

The result is 8.

* What is the average number of rooms (result of running mean() function) in the above 'rooms' vector? Again, best remove the NAs first.

In order to know the average number of rooms I typed:

```{r}
mean(rooms, na.rm = TRUE)
```

Mean is a function which allows me to find the average number of the vector (). Rooms are the object which I assigned the vector to above. Na.rm means NA remove. TRUE is a logical vector I used to test for objects that are logical, which in this case means that it only showed the numbers and not the NA. If I had written FALSE instead of TRUE the result would have been NA. 
The result is: 2.318182 rooms pr. household.

In order to round the number, I used the function:

```{r}
round(x = 2.318182, digits = 0)
```

Now the result of rooms pr. household is: 2

* What type of data is in the 'rooms' vector? What function helps you determine the answer?

To solve this exercise, I typed:

```{r}
class(rooms)
```

This function askes RStudio which type of data is in the () – in this case the object named rooms. 
The result: numeric.
 
* Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object (see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section). Take a screenshot of your RStudio interface showing a) the script you used to create the object, b) the 'interviews' object in the Environment and the c) structure of your R project in the bottom right Files pane. Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS). In this Blackboard submission, list the URL leading to the screenshot in your github repository.

Link to my repository: https://github.com/Digital-Methods-HASS/auau565919_Kendon_Mette


#Final Thoughts
My overall thought on this assignment was that it was not too difficult nor too easy. The R-tutorial was a great help! After completing the exercises, I have a much better idea of how to use RStudio in general but also in regard to my final project. I think that RStudio is perfect for my dataset because it allows you to interact with your dataset in a lot of different ways, whilst saving your steps as a script. 
Besides this, I want to use Overleaf for my final project, but I am struggling a bit with the format. Some of the characters I used for this assignment did not go well in Overleaf and therefor I had to move my learning journal into Word. I hope that I can get a handle on how to use Overleaf before the final project, because I would like to use it. 

